---
title: "p8105_hw6_cz2544"
author: "Chunxiao Zhai"
date: "11/16/2018"
output: github_document
---

```{r setup, include=FALSE}
library(p8105.datasets)
library(tidyverse)
library(ggridges)
library(pander)
library(modelr)
library(mgcv)
panderOptions('round', 3)
panderOptions('keep.trailing.zeros', TRUE)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "95%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

##Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here.

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. 

Modifiy  victim_race to have categories white and non-white, with white as the reference category. Be sure that victim_age is numeric.

```{r load_homicide}
homi_data = read.csv("homicide-data.csv") %>% 
  mutate(city_state = str_c(city, state, sep = ", "),
         solve = if_else(disposition == "Closed by arrest", 1, 0)) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) %>% 
  mutate(victim_race_bin = if_else(victim_race == "White", "white", "non-white"),
         victim_race_bin = relevel(as.factor(victim_race_bin), ref = "white"),
         victim_age = as.numeric(victim_age)) 
```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r glm_logi_Balti}
fit_logi = homi_data %>%
  filter(city == "Baltimore") %>% 
  glm(solve ~ victim_age + victim_race_bin + victim_sex, data = ., family = binomial()) 

OR_data = fit_logi %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         OR_ci_low = exp(pull(broom::confint_tidy(fit_logi), 1)), 
         OR_ci_high = exp(pull(broom::confint_tidy(fit_logi), 2))) %>% 
  select(term, log_OR = estimate, OR, OR_ci_low, OR_ci_high, p.value)

```

With all other variables fixed, the estimate of adjusted odds ratio for solving homicides comparing non-white victims to white victims is `r pull(OR_data, OR)[3]`, the 95% confidence interval is (`r pull(OR_data, OR_ci_low)[3]`, `r pull(OR_data, OR_ci_high)[3]`).

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. Do this within a “tidy” pipeline, making use of  purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r orci_eachcity, warning=FALSE}
homi_data_orci = homi_data %>% 
  group_by(city_state) %>% 
  nest(victim_age, victim_race_bin, victim_sex, solve ) %>% 
  mutate(models_glm = map(data, ~glm(solve ~ victim_age + victim_race_bin + victim_sex, data = ., family = binomial())),
         models = map(models_glm, broom::tidy),
         ci = map(models_glm, broom::confint_tidy)) %>% 
  select(-data, -models_glm) %>% 
  unnest() %>% 
  filter(term  == "victim_race_binnon-white") %>% 
  mutate(OR = exp(estimate),
         OR_ci_low = exp(conf.low), 
         OR_ci_high = exp(conf.high)) %>% 
  select(city_state, starts_with("OR"))
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r plot_orci}
homi_data_orci_plot = 
  homi_data_orci %>% 
  arrange(OR) %>% 
  mutate(city_state = fct_inorder(city_state)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = OR_ci_low, ymax = OR_ci_high), width = 0.2) +
    labs(
    title = "Adjusted OR for solving homicides non-white vs white for 47 large U.S. cities",
    y = "OR estimates",
    caption = "Data from the Washington Post"
  )  +
  theme(axis.text.x = element_text(angle =  -60,size = 6, vjust = -1.4 ))

homi_data_orci_plot
```

Comment: Among the 47 cities, only 3 have ORs slightly over 1 (`r pull(tail(homi_data_orci%>%arrange(OR),3),1)`), probably due to large error. In all other 44 cities, ORs are less than 1, indicating cases of non-white victims are less likely to solve. But in some cities with low OR, the errors are big and the 95% confidence interval contains 1, like Richmand, Washingto and New York. There are `r sum(homi_data_orci$OR_ci_high >= 1)` cities with OR confidence interval contain 1, and `r sum(homi_data_orci$OR_ci_high < 1)` cities with OR confidence interval do not contain 1. Top 5 cities with lowest ORs are `r pull(head(homi_data_orci%>%arrange(OR),5),1)`.


##Problem 2

In this probelm, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, consists of roughly 4000 children and includes the following variables:

babysex: baby’s sex (male = 1, female = 2)
bhead: baby’s head circumference at birth (centimeters)
blength: baby’s length at birth (centimeteres)
bwt: baby’s birth weight (grams)
delwt: mother’s weight at delivery (pounds)
fincome: family monthly income (in hundreds, rounded)
frace: father’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
gaweeks: gestational age in weeks
malform: presence of malformations that could affect weight (0 = absent, 1 = present)
menarche: mother’s age at menarche (years)
mheigth: mother’s height (inches)
momage: mother’s age at delivery (years)
mrace: mother’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
parity: number of live births prior to this pregnancy
pnumlbw: previous number of low birth weight babies
pnumgsa: number of prior small for gestational age babies
ppbmi: mother’s pre-pregnancy BMI
ppwt: mother’s pre-pregnancy weight (pounds)
smoken: average number of cigarettes smoked per day during pregnancy
wtgain: mother’s weight gain during pregnancy (pounds)

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r load_birthweight}
bthwt_data = read.csv("birthweight.csv") 
#look at the variables
skimr::skim(bthwt_data)
#first impression of correlation
round(cor(bthwt_data),3)

#tidy
bthwt_data = bthwt_data %>% 
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         mrace = as.factor(mrace)) 

skimr::skim(bthwt_data)

#look at distribution
ggplot(data = bthwt_data, aes(x = bwt)) + geom_histogram(bins = 100)

#look at outlier of key variables
ggplot(data = bthwt_data, aes(x = bwt, y = gaweeks)) + geom_point()
ggplot(data = bthwt_data, aes(x = bwt, y = delwt)) + geom_point()
ggplot(data = bthwt_data, aes(x = bwt, y = wtgain)) + geom_point()
```

Comment: There is `r sum(is.na(bthwt_data))` missing data in this dataset. There are `r sum(bthwt_data$babysex == 1)` boys and `r sum(bthwt_data$babysex == 2)` girls. Baby’s birth weight is highly correlated with bhead(baby’s head circumference at birth (centimeters)) and 
blength (baby’s length at birth (centimeteres)), which are different dimentions of measurement of the babies' growth. While gaweeks (gestational age in weeks), wtgain (mother’s weight gain during pregnancy (pounds)), delwt (mother’s weight at delivery (pounds)) and mrace (mother’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)) showed correlation greater than 0.2. Among these promising variables, wtgain, delwt are all indicators of mother's weight condition, which will be related to mother's body features like height and bmi, and also related to the term of pregnancy. Mother's race is highly correlated with father's race, but frace is not that much related to baby birth weight, frace is likely to be a confounder.

There are several data points with gestational weeks less than 25 weeks seem far away from the rest of the data points, might deserve special investigation.

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and  add_residuals in making this plot.

```{r bthwt_lm1}
fit1 = lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks + malform + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi + ppwt + smoken + wtgain , data = bthwt_data)

summary(fit1)
```

Comment: When put all variables in the model as predictors, 8 showed high statistical significance at level of 0.05: babysex2, bhead, blength, delwt, gaweeks, mrace2, parity and smoken. fincome has a p-value of 0.1. Considering frace a possible confounder, fit the model again without frace:

```{r bthwt_lm2}
fit2 = lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + malform + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi + ppwt + smoken + wtgain , data = bthwt_data)

summary(fit2)
```

Comment: mrace 3 become significant at 0.1 level but adjusted R squar barely changed, indicating frace contributing nothing to the model. malform and ppbmi have large p-values, fit new model without them: 

```{r bthwt_lm3}
fit3 = lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga  + ppwt + smoken + wtgain , data = bthwt_data)

summary(fit3)
```

Comment: ppwt, mrace 4, mheight become significant, indicating ppbmi is related to these three variables, considering bmi is a better indicator combing weight and height, put ppbmi back and remove ppwt and mheight:

```{r bthwt_lm4}
fit4 = lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + menarche + momage + mrace + parity + pnumlbw + pnumsga  + ppbmi + smoken + wtgain , data = bthwt_data)

summary(fit4)
```

Comment: fincome, mrace3 and wtgain become significant at level of 0.1, momage and menarche are still insignificant. Considering menarche is less likely to relate to baby weight in biological mechanisms, remove it first. When menarche is removed, momage is even less significant. The pnumlbw and pnumsga are not defined because of singularities, by reviewing dataset, we fought these two variables are all 0 for all patients, thus they are removed from model, although highly medically meaningful in concept. 

```{r bthwt_lm5}
fit5 = lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + mrace + parity  + ppbmi + smoken + wtgain , data = bthwt_data)

summary(fit5)
```


Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

```{r bthwt_cv_rmse, warning=FALSE}
set.seed(1)

cv_df =
  crossv_mc(bthwt_data, 4342) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble)) %>% 
  mutate(meo_mod = map(train, ~lm(bwt ~ gaweeks, data = .x)),
         thr_mod = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex 
                                  + bhead*blength*babysex, data = .x)),
         fit5_mod = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + 
                                     mrace + parity  + ppbmi + smoken + wtgain, data = as_tibble(.x)))) %>% 
  mutate(rmse_meo  = map2_dbl(meo_mod, test, ~rmse(model = .x, data = .y)),
         rmse_thr  = map2_dbl(thr_mod, test, ~rmse(model = .x, data = .y)),
         rmse_fit5 = map2_dbl(fit5_mod, test, ~rmse(model = .x, data = .y))) %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

Comment: The main effects only model have much higher rmse than the three predictor interaction model and multi-predictor model. The rmse of three predictor interaction model is slightly higher than the multi-predictor model, but the difference is small.

